---
title: "Exercises: R Basics"
author: Daniel Dauber
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Exercises section of R4NP: A Guide for Social Scientists"
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)

knitr::opts_chunk$set(echo = FALSE)
```

## Knowledge Check

###

```{r q1-order-of-operations, echo=FALSE}
question("What does the following code return: `4 + 3 * 2`?",
         answer("14", message = "Check the order of operations. Multiplication happens before addition."),
         answer("10", correct = TRUE, message = "Correct! 3 * 2 = 6, then 4 + 6 = 10."),
         answer("7", message = "That would be 4 + 3, then * 2 — but R uses standard math order."),
         answer("6", message = "Nope — think carefully about multiplication vs addition."),
         allow_retry = TRUE
)
```

###

```{r q2-operators, echo=FALSE}
question("Which of the following are valid arithmetic operators in R?",
         answer("`+`", correct = TRUE, message = "Correct! `+` is for addition."),
         answer("`**`", message = "Not valid in base R."),
         answer("`-`", correct = TRUE, message = "Correct! `-` is for subtraction."),
         answer("`*`", correct = TRUE, message = "Correct! `*` is for multiplication."),
         allow_retry = TRUE
)
```

###

```{r q3-sqrt, echo=FALSE}
question("What is the result of `sqrt(25 - 9)`?",
         answer("4", correct = TRUE, message = "Correct! 25 - 9 = 16, and sqrt(16) = 4."),
         answer("2", message = "Check your subtraction — 25 - 9 = 16, not 4."),
         answer("16", message = "That’s the inside of the sqrt, not the final result."),
         answer("Error", message = "No error here — R can handle square roots."),
         allow_retry = TRUE
)
```

###

```{r q4-compare-string-number, echo=FALSE}
question('What will `5 == 5` return in R?',
         answer("Nothing", message = "R always provides an 'answer'. Have another go at it."),
         answer("TRUE", correct = TRUE, message = "Correct! The `==` symbol stands for 'is equal to'."),
         answer("Error", message = "No error — just a logical comparison."),
         answer('"FALSE"', message = "Both values appear to be the same. So they are euqal. Try again"),
         allow_retry = TRUE
)
```

###

```{r q5-logical-operators, echo=FALSE}
question("Which of the following are logical operators in R?",
         answer("==", correct = TRUE, message = "Correct! It checks equality."),
         answer("!=", correct = TRUE, message = "Correct! It checks inequality."),
         answer("&", correct = TRUE, message = "Correct! It's a logical AND."),
         answer("=", message = "This is for assignment, not comparison."),
         allow_retry = TRUE
)

```

###

```{r q6-object-creation, echo=FALSE}
question("What happens when you run `my_number <- 42`?",
         answer("You print the number 42", message = "It doesn't print unless you explicitly tell it do so."),
         answer("You create an object", correct = TRUE, message = "Yes! `my_number` is now stored in your environment."),
         answer("You assign 42 to `my_number`", correct = TRUE, message = "Correct — that’s exactly what the `<-` does."),
         answer("It returns an error", message = "No error — this is valid R syntax."),
         allow_retry = TRUE
)
```

###

```{r q7-assignment-symbols, echo=FALSE}
question("Which symbol is used for assignment in R?",
         answer("<-", correct = TRUE, message = "Correct! This is the standard assignment symbol in R."),
         answer("=", correct = TRUE, message = "Technically valid — but `<-` is preferred for clarity."),
         answer("->", correct = TRUE, message = "Correct — but it assigns in the opposite direction."),
         answer("=>", message = "That’s from other languages like JavaScript."),
         allow_retry = TRUE
)
```

###

```{r q8-character-vector, echo=FALSE}
question('What does `c("a", "b", "c")` create?',
         answer("A character vector", correct = TRUE, message = "Correct — it combines character elements into a vector."),
         answer("A list", message = "Nope — use `list()` for that."),
         answer("A data frame", message = "A data frame needs named columns and structure."),
         answer("An object", message = "Without `<-` it would not become a new object."),
         allow_retry = TRUE
)
```

###

```{r q9-assignment, echo=FALSE}
question("What happens when you run the following code in RStudio: `x <- 4 + 5`?",
         answer("The result `9` is printed in the console.", message = "Not unless you explicitly call `x`."),
         answer("The result is saved in an object called `x`.", correct = TRUE, message = "Correct — `x` stores the result."),
         answer("An object named `4` is created.", message = "No — object names must come before the assignment arrow."),
         answer("An error is returned.", message = "This is valid syntax — no error."),
         allow_retry = TRUE
)
```

###

```{r q10-naming_conventions, echo=FALSE}
question("Which of the following is a valid object name in R?",
         answer("2nd_name", message = "Object names cannot begin with a number."),
         answer("first_name", correct = TRUE, message = "Correct — it starts with a letter and uses only valid characters."),
         answer("full name", message = "Spaces are not allowed in object names."),
         answer("_data", message = "Object names cannot start with an underscore."),
         allow_retry = TRUE
)
```

###

```{r q11-loading-packages}
question(
  "What are three steps to use a function from a new R package you found on CRAN?",
  answer("Install the R package using 'install.packages()'", correct = TRUE),
  answer("Load the R package with 'library()'", correct = TRUE),
  answer("Call the function of interest.", correct = TRUE),
  answer("Use 'package_name::function_name()' to use functions directly after installation.", correct = TRUE),
  allow_retry = TRUE
)
```

###

```{r q12-search-packages}
question(
  "Where in RStudio can you install R packages?",
  answer("In the console.", correct = TRUE),
  answer("In the 'Environment' pane.", message = "Unfortuntaely not. You can only find objects in the 'Environment' pane."),
  answer("In the 'Packages'.", correct = TRUE),
  answer("In the 'Source' window, using an R script.", correct = TRUE),
  allow_retry = TRUE
)
```

## Coding Practice

### Basic Computations | 01

Compute the result of 7 plus 5.

```{r arithmetic-1, exercise=TRUE}

```

```{r arithmetic-1-hint-1}
# Use the + operator to add the two numbers
```

```{r arithmetic-1-hint-2}
# Example: 2 + 2
```

```{r arithmetic-1-solution}
7 + 5
```

```{r arithmetic-1-check}
grade_this_code()
```

### Basic Computations | 02

Check if 3 is greater than 1.

```{r logical-1, exercise=TRUE}

```

```{r logical-1-hint-1}
# Use the > operator
```

```{r logical-1-hint-2}
# Syntax: 3 > 1
```

```{r logical-1-solution}
3 > 1
```

```{r logical-1-check}
grade_this_code()
```

### Basic Computations | 03

Subtract 8 from 20.

```{r arithmetic-2, exercise=TRUE}

```

```{r arithmetic-2-hint-1}
# Use the - operator
```

```{r arithmetic-2-hint-2}
# Syntax: 20 - 8
```

```{r arithmetic-2-solution}
20 - 8
```

```{r arithmetic-2-check}
grade_this_code()
```

### Basic Computations | 04

Assign the value 42 to an object called 'answer'.

```{r object-1, exercise=TRUE}

```

```{r object-1-hint-1}
# Use <- to assign a value
```

```{r object-1-hint-2}
# Syntax: object_name <- value
```

```{r object-1-solution}
answer <- 42
```

```{r object-1-check}
grade_this_code()
```

### Basic Computations | 05

Please, create a list called `books` and include the following book titles in it in this order:

-   "Harry Potter and the Deathly Hallows",

-   "The Alchemist",

-   "The Davinci Code",

-   "R For Dummies".

Then, inspect the object `books` to see the book titles inside of it.

```{r book-list, exercise=TRUE, exercise.lines=6}
# Create a list of book titles
```

```{r book-list-hint-1}
# Create an object called 'books'
books <- ___
```

```{r book-list-hint-2}
# You need to use the function 'list()' to create a list object
books <- list(___)
```

```{r book-list-hint-3}
# Remember to use "" for each book title
books <- list("Harry Potter and the Deathly Hallows")
```

```{r book-list-solution}
books <- list("Harry Potter and the Deathly Hallows",
              "The Alchemist",
              "The Davinci Code",
              "R For Dummies")
books
```

```{r book-list-check}
grade_this({
  correct <- list(
    "Harry Potter and the Deathly Hallows",
    "The Alchemist",
    "The Davinci Code",
    "R For Dummies"
  )
  user_val <- get("books", envir = .envir_result)

  if (setequal(unlist(user_val), unlist(correct))) {
    pass("Great! Your list contains all the correct book titles.")
  }

  fail("It looks like the `books` list is missing titles or has incorrect entries.")
})
```

### Basic Computations | 06

Create an object `todo` and concatenate (`c()`) the following three tasks:

-   Go shopping

-   Prepare lunch

-   Have coffee with Peter Parker

```{r todo, exercise=TRUE}
# Create a todo list
```

```{r todo-hint-1}
# Create an object 'todo'
todo <- ___
```

```{r todo-hint-2}
# Use the function 'c()'
todo <- c(___)
```

```{r todo-solution}
todo <- c("Go shopping",
          "Prepare lunch",
          "Have coffee with Peter Parker")
```

```{r todo-check}
grade_this({
  correct <- c("Go shopping", "Prepare lunch", "Have coffee with Peter Parker")
  user_val <- get("todo", envir = .envir_result)

  if (setequal(user_val, correct)) {
    pass("Excellent! You've listed all the right tasks.")
  }

  fail("Double-check your `todo` list — something's missing or incorrect.")
})
```

### Basic Computations | 07

Can you find out which cookie is my favourite one in `file_daniel`? Provide the code that returns my favourite cookie.

```{r nested-list-setup, include=FALSE}
# Movies
movies <- c("Captain America; The first Avenger",
            "Lala Land",
            "Moana",
            "Dr Strange")

# Food
savory <- c("Laksa Noodles",
            "Pan con tomate",
            "Shanghainese Maifun")

cakes <- c("NY-style cheesecake",
           "Burnt cheesecake",
           "White chocolate oolong sponge cake")

cookies <- c("White chocolate macademia cookie") 

food <- list(savory, cakes, cookies)
names(food) <- c("Savory-food", "Cakes", "Cookies")

file_daniel <- list(movies, food)
names(file_daniel) <- c("Movies", "Food")
```

```{r nested-list, exercise=TRUE, exercise.setup = "nested-list-setup"}
# Find out which cookie is Daniel's favourite
```

```{r nested-list-hint-1}
# Use $ after the object name to dig into the list
```

```{r nested-list-hint-2}
file_daniel$Food$
```

```{r nested-list-hint-3}
file_daniel$Food$Cookies
```

```{r nested-list-hint-4}
file_daniel
```

```{r nested-list-solution}
file_daniel$Food$Cookies
```

```{r nested-list-check}
grade_this_code()
```

### Functions | 01

```{r function-1, exercise=TRUE}
# Find the length of the vector c(2, 4, 6, 8)
```

```{r function-1-hint-1}
# Use the length() function
```

```{r function-1-hint-2}
# Syntax: length(vector)
```

```{r function-1-solution}
length(c(2, 4, 6, 8))
```

```{r function-1-check}
grade_this_code()
```

### Functions | 02

```{r function-2, exercise=TRUE}
# Use abs() to find the absolute value of -15
```

```{r function-2-hint-1}
# Use abs() function
```

```{r function-2-hint-2}
# Syntax: abs(number)
```

```{r function-2-solution}
abs(-15)
```

```{r function-2-check}
grade_this_code()
```

### Functions | 03

Create a vector named `cookie_sales` that contains the following sales numbers and compute the average: 55, 76, 32, 49.

```{r function-3, exercise=TRUE}

```

```{r function-3-hint-1}
# Use c() function to combine all numbers into one vector.
```

```{r function-3-hint-2}
# The 'average' is called 'mean' in R.
```

```{r function-3-solution}
cookie_sales <- c(55, 76, 32, 49)

mean(cookie_sales)
```

```{r function-3-check}
grade_this_code()
```

### Coding Etiquette | 01

Assign the number 4 to an object called `my_fav_number` and include the comment `My lucky number`.

```{r etiquette-1, exercise=TRUE}

```

```{r etiquette-1-hint-1}
# Use <- for assignment and # for comments
```

```{r etiquette-1-hint-2}
# Example: my_var <- 5  # this is a comment
```

```{r etiquette-1-solution}
my_favorite_number <- 4  # My lucky number
```

```{r etiquette-1-check}
grade_this_code()
```

### Coding Etiquette | 02

Create an object called `my_name` and assign your name to it (as a character string).

```{r etiquette-2, exercise=TRUE}

```

```{r etiquette-2-hint-1}
# Use quotation marks around your name
```

```{r etiquette-2-hint-2}
# Example: my_name <- "Alice"
```

```{r etiquette-2-solution}
my_name <- "Alice"
```

```{r etiquette-2-check}
grade_this_code()
```

### Coding Etiquette | 03

Make this code chunk more readable. Be aware, there are many different ways to do this. The solution provided is just one of many.

```{r etiquette-3, exercise=TRUE}
X=5+7#add apples
MyDATA<-list("Apple",TRUE,  7.1);print(MyDATA)
```

```{r etiquette-3-hint-1}
# Use clear naming and spacing
```

```{r etiquette-3-hint-2}
# Example: number_of_apples <- 5 + 7
```

```{r etiquette-3-solution}
# Add the number of apples collected
apple_count <- 5 + 7

# Create a list with fruit name, freshness, and weight
my_data <- list("Apple", TRUE, 7.1)

# Print the list to inspect its contents
print(my_data)
```


## Case Study Challenge

### Packing for a Weekend Trip | 01

You are preparing for a weekend trip to Paris. Over the last weeks you prepared a packing list which you stored in an object called `packing_list`. 

Your task is to create a new object called `already_packed` that includes the items you’ve already packed: `"shirt"`, `"toothbrush"`, and `"socks"`.

```{r packing-setup, include=FALSE}
packing_list <- c("shirt", "trousers", "toothbrush", "passport", "socks")
```

```{r already-packed, exercise=TRUE, exercise.setup="packing-setup"}

```

```{r already-packed-hint-1}
# Use c() to create a character vector
```

```{r already-packed-hint-2}
# Your list should contain three items: shirt, toothbrush, and socks
```

```{r already-packed-solution}
already_packed <- c("shirt", "toothbrush", "socks")
```

```{r already-packed-check}
grade_this_code()
```

### Packing for a Weekend Trip | 02

Use the function `length()` to find out how many items are on the full `packing_list`.

```{r packing-length, exercise=TRUE, exercise.setup="packing-setup"}

```

```{r packing-length-hint-1}
# Use the length() function on packing_list
```

```{r packing-length-hint-2}
# Syntax: length(object_name)
```

```{r packing-length-solution}
length(packing_list)
```

```{r packing-length-check}
grade_this_code()
```

### Packing for a Weekend Trip | 03

Now use `length()` again to find out how many items you have already packed (using your `already_packed` object).

```{r already-packed-length-setup, include=FALSE}
already_packed <- c("shirt", "toothbrush", "socks")
```

```{r already-packed-length, exercise=TRUE, exercise.setup="already-packed-length-setup"}

```

```{r already-packed-length-hint-1}
# Use the length() function on your object
```

```{r already-packed-length-hint-2}
# Syntax: length(already_packed)
```

```{r already-packed-length-solution}
length(already_packed)
```

```{r already-packed-length-check}
grade_this_code()
```

### Packing for a Weekend Trip | 04

Now inspect both `packing_list` and `already_packed` to figure out what item is missing. Create an object named `forgotten_item` and assign the missing item to it as a character string. Include the code for inspecting the `packing_list` and the `already_packed` object.

```{r forgotten-item-setup, include=FALSE}
packing_list <- c("shirt", "trousers", "toothbrush", "passport", "socks")
already_packed <- c("shirt", "toothbrush", "socks")
```

```{r forgotten-item, exercise=TRUE, exercise.setup="forgotten-item-setup"}

```

```{r forgotten-item-hint-1}
# Use quotation marks and assign the value using <-
```

```{r forgotten-item-hint-2}
# Think: what’s in packing_list that’s not in already_packed?
```

```{r forgotten-item-solution}
packing_list
already_packed

forgotten_item <- c("trousers", "passport")
```

```{r forgotten-item-check}
grade_this({
  if (!exists("forgotten_item", envir = .envir_result)) {
    fail("It looks like you didn't assign an object called `forgotten_item`.")
  }

  correct <- c("trousers", "passport")
  user_val <- get("forgotten_item", envir = .envir_result)

  if (setequal(user_val, correct)) {
    pass("Excellent! You correctly assigned the forgotten items to `forgotten_item`.")
  }

  fail("Almost. Make sure `forgotten_item` includes all the missing items.")
})
```
