---
title: "Exercises: R Basics"
author: Daniel Dauber
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Exercises section of R4NP: A Guide for Social Scientists"
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.blanks = TRUE)
```

## Basic computations

### Exercise 1

Change the code below to divide the result of `(2022-1982)` by `5`.

```{r division, exercise=TRUE}
(2022-1982)
```

```{r division-solution}
(2022-1982)/5
```

### Exercise 2

What is the result of $\sqrt[2]{25-16}+2*8-6$?

```{r take-sqrt, exercise=TRUE}

```

```{r take-sqrt-hint-1}
# You need a function to compute the square root
```

```{r take-sqrt-hint-2}
sqrt(___)
```

```{r take-sqrt-solution}
sqrt(25-16)+2*8-6
```

### Exercise 3

What does the console return if you execute the following code `"Five" == 5`?

```{r logicals, exercise=TRUE}

```

```{r logicals-quiz,}
quiz(
  question("What does the result imply?",
           answer("'Five' and '5' are the same."),
           answer("'Five' and '5' are not the same.", correct = TRUE),
           answer("We performed a logical operation.", correct = TRUE),
           allow_retry = TRUE
           )
)
```

## Assigning values to objects: `<-`

### Exercise 4

First, create a list called `books` and include the following book titles in it:

-   "Harry Potter and the Deathly Hallows",

-   "The Alchemist",

-   "The Davinci Code",

-   "R For Dummies".

Secondly, inspect the object `books` to see the book titles inside of it.

```{r book-list, exercise=TRUE, exercise.lines=6}

```

```{r book-list-hint-1}
# Create an object called 'books'
books <- ___
```

```{r book-list-hint-2}
# You need to use the function 'list()' to create a list object
books <- list(___)
```

```{r book-list-hint-3}
# Remember that you have to use "" for each book title, e.g.
books <- list("Harry Potter and the Deathly Hallows")
```

```{r book-list-solution}
# The solution
books <- list("Harry Potter and the Deathly Hallows",
              "The Alchemist",
              "The Davinci Code",
              "R For Dummies")

# Inspect the list of `books`
books
```

### Exercise 5

Create an object `todo` and concatenate (`c()`) the following three tasks:

-   Go shopping

-   Prepare lunch

-   Have coffee with Peter Parker

```{r todo, exercise=TRUE}

```

```{r todo-hint-1}
# Create an object 'todo'
todo <- ___
```

```{r todo-hint-2}
# Use the function 'c()'
todo <- c(___)
```

```{r todo-solution}
# The solution
todo <- c("Go shopping",
          "Prepare lunch",
          "Have coffee with Peter Parker")
```

### Exercise 6

Can you find out which cookie is my favourite one in `file_daniel`?

```{r nested-list-setup, include=FALSE}
# Movies
movies <- c("Captain America; The first Avenger",
            "Lala Land",
            "Moana",
            "Dr Strange")

# Food
savory <- c("Laksa Noodles",
            "Pan con tomate",
            "Shanghainese Maifun")

cakes <- c("NY-style cheesecake",
           "Burnt cheesecake",
           "White chocolate oolong sponge cake")

cookies <- c("White chocolate macademia cookie") 

food <- list(savory,
             cakes,
             cookies)

names(food) <- c("Savory-food", "Cakes", "Cookies")

# Compile file
file_daniel <- list(movies,
             food)

names(file_daniel) <- c("Movies", "Food")
```

```{r nested-list, exercise=TRUE, exercise.setup = "nested-list-setup"}

```

```{r nested-list-hint-1}
# Option 1: You use '$' after the object name to dig deeper into the list
```

```{r nested-list-hint-2}
# Option 1: You use '$' after the object name to dig deeper into the list
file_daniel$Food$
```

```{r nested-list-hint-3}
# Option 1: You use '$' after the object name to dig deeper into the list
file_daniel$Food$Cookies
```

```{r nested-list-hint-4}
# Option 2: You inspect the entire list in the console
file_daniel
```

## Functions

### Exercise 7

Use the following function and give it the following two values: `2` and `3`.

```{r function, exercise=TRUE, exercise.lines = 8}
x_x <- function(number1, number2){
      result1 <- number1 * number2
      result2 <- sqrt(number1)
      result3 <- number1 - number2
      return(c(result1, result2, result3))
    }
```

```{r function-hint-1}
# To use the function you have to write and insert the values.
x_x(___)
```

```{r function-hint-2}
# You can insert the values as is, because they are numbers.
x_x(2, 3)
```

```{r function-quiz}
quiz(
  question(
    "What does the computation in the function do?",
    answer("It adds up the numbers '2' and '3'.",
           message = "Almost. The function subtracts first value from the second."),
    answer("It takes the squareroot of the second number.",
           message = "Not quite. It takes the squareroot of the first number."),
    answer("It multiplies the first number with the second number", correct = TRUE),
    allow_retry = TRUE
  )
)
```

## R packages

### Exercise 8

```{r loading-packages}
quiz(
  question(
    "What are three steps to use a function from a new R package you found on CRAN?",
    answer("Install the R package using 'install.packages()'", correct = TRUE),
    answer("Load the R package with 'library()'", correct = TRUE),
    answer("Call the function of interest.", correct = TRUE),
    answer("Use 'package_name::function_name()' to use functions directly after installation.", correct = TRUE),
    allow_retry = TRUE
  )
)
```

```{r search-packages}
quiz(
  question(
    "Where in RStudio can you install R packages?",
    answer("In the console.", correct = TRUE),
    answer("In the 'Environment' pane.", message = "Unfortuntaely not. You can only find objects in the 'Environment' pane."),
    answer("In the 'Packages'.", correct = TRUE),
    answer("In the 'Source' window, using an R script.", correct = TRUE),
    allow_retry = TRUE
  )
)

```
