---
title: "Exercises: Correlations"
author: Daniel Dauber
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
    toc: true
    toc_depth: 3
runtime: shiny_prerendered
description: "Exercises section of R for Non-Programmers (R4NP)"
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(r4np)
library(tidyverse)
library(correlation)
knitr::opts_chunk$set(echo = FALSE)
```

## Knowledge Checks

### Knowledge Check | 1
```{r corr-q1, echo=FALSE}
question("What does a correlation coefficient measure?",
  answer("The difference in means between groups.",
         message = "That's for t-tests or ANOVA, not correlation. We will learn more about in the next chapter."),
  answer("The strength and direction of association between two variables.", correct = TRUE,
         message = "Correct! Correlation quantifies association, not causation."),
  answer("The probability that results are due to chance.",
         message = "That's a p-value, not a correlation coefficient."),
  answer("How much two variables cause each other.",
         message = "Correlation doesn't imply causation."),
  allow_retry = TRUE
)
```

### Knowledge Check | 2
```{r corr-q2, echo=FALSE}
question("What is the possible range for a Pearson correlation coefficient?",
  answer("0 to 1",
         message = "Almost, correlations can be negative as well as positive."),
  answer("-100 to 100",
         message = "No quite, correlation coefficients are never this large."),
  answer("0 to infinity",
         message = "Sorr, correlations are always between -1 and 1."),
  answer("From -1 to 1", correct = TRUE,
         message = "That's right! Correlations range from -1 to 1."),
  allow_retry = TRUE
)
```

### Knowledge Check | 3
```{r corr-q3, echo=FALSE}
question("If two variables have a strong correlation, can we conclude that one causes the other?",
  answer("Yes, if r > 0.7.",
         message = "Even a high r does not guarantee causation."),
  answer("Yes, but only if p < 0.05.",
         message = "Statistical significance is not causation."),
  answer("Yes, always.",
         message = "I am afraid, we never can assume causation from correlation."),
  answer("No, correlation does not imply causation.", correct = TRUE,
         message = "Exactly! Correlation is not causation."),
  allow_retry = TRUE
)
```

### Knowledge Check | 4
```{r corr-q4, echo=FALSE}
question("Which of these r values shows the strongest relationship?",
  answer("-0.82", correct = TRUE,
         message = "Right! The sign does not matter for strength—only the magnitude."),
  answer("0.60",
         message = "This is a moderate positive relationship."),
  answer("0.04",
         message = "This is a very weak association."),
  answer("0.33",
         message = "This is a weak positive association."),
  allow_retry = TRUE
)
```

### Knowledge Check | 5
```{r corr-q5, echo=FALSE}
question("What does partial correlation allow us to do?",
  answer("Test if variables are normally distributed.",
         message = "That's not what partial correlation does."),
  answer("Measure the association between two variables while controlling for a third.", correct = TRUE,
         message = "Correct! This is the unique strength of partial correlation."),
  answer("Increase sample size.",
         message = "Partial correlation doesn't change your sample size."),
  answer("Test the difference in means between two groups.",
         message = "That's for t-tests, not correlation."),
  allow_retry = TRUE
)
```

## Coding Practice

### Coding Practice | 1a

A research team is exploring whether time spent studying predicts exam scores among university students. The dataset `study_data` provide information about `study_hours` and `exam_score`. Visualise their relationship using a scatterplot. Plot `study_hours` on to the x-axis.

```{r cp1-corr-plot-setup, include=FALSE}
set.seed(10)
study_data <- tibble::tibble(
  study_hours = round(rnorm(100, 7, 2), 1),
  exam_score = round(65 + 3 * study_hours + rnorm(100, 0, 10), 1)
)
```

```{r cp1-corr-plot, exercise=TRUE, exercise.setup="cp1-corr-plot-setup"}
# Use ggplot and geom_point() to show the association.
```

```{r cp1-corr-plot-hint}
# Use ggplot2 and geom_point().
```

```{r cp1-corr-plot-solution}
study_data |> 
  ggplot(aes(x = study_hours, y = exam_score)) +
  geom_point()
```

```{r cp1-corr-plot-check}
grade_this_code()
```

### Coding Practice | 1b

```{r cp1-corr-plot-mcq, echo=FALSE}
question("What does the scatterplot suggest?",
  answer("There is a negative association.",
         message = "I am afraid, that's not what the plot shows."),
  answer("There is no relationship at all.",
         message = "Look at the trend in the points."),
  answer("There is a positive association: more study hours are linked with higher scores.", correct = TRUE,
         message = "Correct! The points trend upward."),
  answer("Most students studied around 7 hours.",
         message = "The data show a range of study hours and seem quite nicely spread out."),
  allow_retry = TRUE
)
```

### Coding Practice | 2a

A survey (`coffee`) asked 80 students how many cups of coffee they drink each day and their final grade in an online class.

However, the plot looks rather odd. Although 80 responses were collected, it seems much less are shown in the plot. Have a look at the plot first and consider what the potential problem could be. Then consider how to amend the code for a better visualisation of relationships.

```{r cp2-jitter-setup, include=FALSE}
set.seed(123)
cups <- sample(0:5, 80, replace = TRUE)
grades <- sample(seq(50, 100, 5), 80, replace = TRUE)
coffee <- tibble::tibble(
  cups = cups,
  grade = grades
)
```

```{r cp2-jitter-plot, exercise=TRUE, exercise.setup="cp2-jitter-setup"}
# Plot cups vs. grade so that overlapping points are revealed.
coffee |> 
  ggplot(aes(x = cups, y = grade)) +
  geom_point(width = 0.25, height = 0.5)
```

```{r cp2-jitter-plot-hint}
# Use geom_jitter() instead of geom_point().
```

```{r cp2-jitter-plot-solution}
coffee |> 
  ggplot(aes(x = cups, y = grade)) +
  geom_jitter(width = 0.25, height = 0.5)
```

```{r cp2-jitter-plot-check}
grade_this_code()
```

### Coding Practice | 2b

```{r cp2-jitter-plot-mcq, echo=FALSE}
question("Why is geom_jitter() useful here?",
  answer("It makes the data points smaller.",
         message = "No, jitter just changes their position, not their size."),
  answer("It spreads out overlapping data so we can see each point.", correct = TRUE,
         message = "Correct! Jittering reveals hidden points."),
  answer("It sorts the data by value.",
         message = "Jitter doesn't sort, it spreads points randomly."),
  answer("It colors the points by grade.",
         message = "Jitter does not change color unless told to."),
  allow_retry = TRUE
)
```

### Coding Practice | 3a

Researchers want to test if there’s a significant correlation between weekly `exercise` hours and `happiness` scores in adults. Calculate the Pearson correlation coefficient using the `exercise_data` .

```{r cp3-corrtest-setup, include=FALSE}
set.seed(99)
exercise_data <- tibble::tibble(
  exercise = round(rnorm(60, 3, 1.2), 1),
  happiness = round(50 + 7 * exercise + rnorm(60, 0, 10), 1)
)
```

```{r cp3-corrtest, exercise=TRUE, exercise.setup="cp3-corrtest-setup"}
# Use the correlation() function to test the correlation between exercise and happiness.
```

```{r cp3-corrtest-hint}
# Use correlation::correlation() and supply the two columns as a data frame.
```

```{r cp3-corrtest-solution}
library(correlation)
exercise_data |> correlation::correlation()
```

```{r cp3-corrtest-check}
grade_this({
  code_text <- paste(deparse(.user_code), collapse = " ")
  if (!grepl("correlation::correlation|correlation\\(", code_text)) {
    fail("Did you use the `correlation()` function from the correlation package?")
  }
  pass("Excellent! You calculated the correlation using the correlation package.")
})
```

### Coding Practice | 3b

```{r cp3-corrtest-mcq-1, echo=FALSE}
question("Is the correlation statistically significant?",
  answer("Yes, the p-value is small, so the association is significant.", correct = TRUE,
         message = "Spot on! The p-value is less than 0.05."),
  answer("No, the p-value is large.",
         message = "Check the output of the correlation() function."),
  answer("No, there is no correlation.",
         message = "Check the test output for the actual result."),
  answer("The test could not be computed.",
         message = "The correlation() function ran fine. Please, check your code and output in the console."),
  allow_retry = TRUE
)
```

### Coding Practice | 3c

```{r cp3-corrtest-mcq-2, echo=FALSE}
question("How would you rate the strength of the association?",
  answer("Quite strong, with an `r > 0.7`.", correct = TRUE,
         message = "That is right! The effect size is quite substantial."),
  answer("The correlation shows a mild relationship between both variables.",
         message = "The `r` value is actually quite high."),
  answer("There is a strong positive relationship.", correct = TRUE,
         message = "Very good."),
  answer("The test show a non-significant result. Therefore, even a high `r` value is not a reliable measure of the relationship.",
         message = "Not quite right. The `p` value is below 0.05. This means the result is significant."),
  allow_retry = TRUE
)
```

## Case Study | The missing relationship

A school district is examining the relationship between students’ time spent studying and their exam scores and collected data shown in the `study_school` dataframe. Surprisingly, looking at the overall sample, there seems to be no clear correlation between `study_hours` and `exam_score`. But perhaps something is masking this link?

Let's investigate—first, by visualising the basic relationship. Plot `study_hours` on the x-axis. 

```{r cp4-hidden-link-setup, include=FALSE}
set.seed(2025)
n <- 100

school <- as_factor(rep(c("East High", "West High"), each = n/2))

# East High: low hours, high scores
east_hours <- round(rnorm(n/2, mean = 4, sd = 1), 1)
east_scores <- round(70 + 2.5 * east_hours + rnorm(n/2, 0, 4), 1)

# West High: high hours, low scores
west_hours <- round(rnorm(n/2, mean = 6, sd = 1), 1)
west_scores <- round(60 + 2.5 * west_hours + rnorm(n/2, 0, 4), 1)

study_hours <- c(east_hours, west_hours)
exam_score <- c(east_scores, west_scores)

study_school <- tibble::tibble(
  school = school,
  study_hours = study_hours,
  exam_score = exam_score
)
study_school$study_hours <- pmax(0, study_school$study_hours)
```

### Step 1a: Visualise the basic relationship

```{r cp4-hidden-link-plot, exercise=TRUE, exercise.setup="cp4-hidden-link-setup"}
# Plot study hours against exam score for all students
```

```{r cp4-hidden-link-plot-hint}
# Use ggplot2 and geom_point().
```

```{r cp4-hidden-link-plot-solution}
study_school |> 
  ggplot(aes(x = study_hours, y = exam_score)) +
  geom_point()
```

```{r cp4-hidden-link-plot-check}
grade_this_code()
```

### Step 1b: Visualise the basic relationship

```{r cp4-hidden-link-mcq, echo=FALSE}
question("What does this plot suggest about the link between study hours and exam scores?",
  answer("There is a very strong positive correlation.",
         message = "Not quite. It's hard to see a strong trend."),
  answer("There is a strong negative correlation.",
         message = "I am afraid, that's not what the scatterplot shows."),
  answer("There is no obvious correlation overall.", correct = TRUE,
         message = "Right! The pattern is washed out when combining all students."),
  answer("All students studied the same number of hours.",
         message = "Unfortunately that is not right. The data show a spread in study hours."),
  allow_retry = TRUE
)
```

### Step 2a: Calculate the Pearson correlation

It is one thing to plot the data, but maybe by computing the correlation we can obtain a better understanding of the relationship between the two variables. 

```{r cp4-hidden-link-cor, exercise=TRUE, exercise.setup="cp4-hidden-link-setup"}
# Calculate the correlation between study hours and exam scores overall
```

```{r cp4-hidden-link-cor-hint}
# Use correlation() from the correlation package to perform a Pearson correlation.
```

```{r cp4-hidden-link-cor-solution}
library(correlation)
study_school |> correlation()
```

```{r cp4-hidden-link-cor-check}
grade_this_code()
```

### Step 2b: Calculate the Pearson correlation

```{r cp4-hidden-link-cor-mcq, echo=FALSE}
question("Is the correlation between study hours and exam scores significant overall?",
  answer("No, it's not strong, but statistically significant.",
         message = "This is not quite right. Try again."),
  answer("No, it is moderately strong, but not statistically significant.",
         message = "Check your code—was cor.test() used?"),
  answer("No, the correlation is weak and not statistically significant.", correct = TRUE,
         message = "You got it! There is little relationship overall."),
  answer("No, the correlation is almost `0` and not significant.", correct = TRUE,
         message = "This is right."),
  allow_retry = TRUE
)
```

### Step 2c: Calculate the Pearson correlation


```{r cp4-hidden-link-cor-mcq2, echo=FALSE}
question("What could be a way forward to check whether there is truly no relationship given our dataset?",
  answer("We could attempt a partial correlation.", correct = TRUE,
         message = "This is right."),
  answer("Nothing, unless we collect more data.",
         message = "Not quite. We do have one more variable in our dataset."),
  answer("We could rerun our analysis using a different correlation technique.",
         message = "While this is could be an option it would depend on the nature of our data. For this dataset the Pearson correlation is the best choice, though."),
  answer("We could consider other variables in our dataset.", correct = TRUE,
         message = "Yes, we have another variable (`school`) that we could consider."),
  allow_retry = TRUE
)
```

### Step 3a: Uncovering a possible relationship

Let's now check the correlation between study hours and exam scores **controlling for school attended**.

```{r cp4-hidden-link-pcor, exercise=TRUE, exercise.setup="cp4-hidden-link-setup"}

```

```{r cp4-hidden-link-pcor-hint}
# Use group_by() to compute the correlation for each school separately.
```

```{r cp4-hidden-link-pcor-solution}
study_school |>
  group_by(school) |> 
  correlation()
```

```{r cp4-hidden-link-pcor-check}
grade_this_code()
```

### Step 3b: Uncovering a possible relationship

```{r cp4-hidden-link-pcor-mcq, echo=FALSE}
question("What happens when we control for school attended?",
  answer("There is still no significant relationship.",
         message = "Not quite—look at the p-value and effect size."),
  answer("The correlation becomes negative.",
         message = "I am afraid, that's not what the output suggests."),
  answer("The test is invalid.",
         message = "Actually, partial correlation is valid here and useful. Try again."),
  answer("A strong positive correlation emerges between study hours and exam scores.", correct = TRUE,
         message = "Exactly! Controlling for school reveals the real relationship."),
  allow_retry = TRUE
)
```

### Step 4a: Presenting the results to the School District

Now that we gained the important insights it woul be great to visualise the final result. We alread created a scatterplot. We want to recreate it but make sure it shows the right relationships.

```{r cp4-final-plot-to-school, echo=FALSE}
question("What could we do to improve our previous scatterplot?",
  answer("Not really much. We can mention in the title that the correlation is spurious.",
         message = "A possible option, but not ideal given that the plot would not show differences between groups."),
  answer("Change the size of the plot to make dots more visible.",
         message = "Not quite, changing the size of the plot would not reveal the relationships in disguise."),
  answer("Change the shape of the points for each school.", correct = TRUE,
         message = "It is a great idea to change the shape of the points, which would help a little to distinguish the two groups."),
  answer("Change the colour of the dots to indicate which observation belongs to which school", correct = TRUE,
         message = "Using contrasting colours is probably the best option."),
  allow_retry = TRUE
)
```

### Step 4b: Visualising Group Differences in Study Outcomes

While the original scatterplot gives an overall sense of the data, it hides the key insight—differences between schools. By assigning both color and shape to the `school` variable, we make group patterns much more visible. This approach allows stakeholders in the school district to instantly spot hidden trends.

In order to achieve this we can change add `colour` and `shape` within the `aes()` function. Complete the code shown.

```{r cp4-final-plot, exercise=TRUE, exercise.setup="cp4-hidden-link-setup"}
study_school |>
  ggplot(aes(x = study_hours, y = exam_score)) +
  geom_point(size = 2) +
  labs(
    title = "Study Hours vs Exam Scores by School",
    x = "Study Hours per Week",
    y = "Exam Score"
  )
```

```{r cp4-final-plot-hint}
# In ggplot, add both `color = school` and `shape = school` in aes(). It is also fine to use `col = school` or `colour = school`.
```

```{r cp4-final-plot-solution}
study_school |>
  ggplot(aes(x = study_hours, y = exam_score, color = school, shape = school)) +
  geom_point(size = 2) +
  labs(
    title = "Study Hours vs Exam Scores by School",
    x = "Study Hours per Week",
    y = "Exam Score"
  )
```

```{r cp4-final-plot-check}
grade_this({
  code_text <- paste(deparse(.user_code), collapse = " ")
  # Accept color, colour, or col for color aesthetic
  if (!grepl("(color|colour|col)\\s*=\\s*school", code_text)) {
    fail("Did you map `color`, `colour`, or `col` to `school` in your plot?")
  }
  # Accept both shape = school or pch = school
  if (!grepl("(shape|pch)\\s*=\\s*school", code_text)) {
    fail("Did you map `shape` or `pch` to `school` in your plot?")
  }
  if (!grepl("geom_point", code_text)) {
    fail("Make sure to use `geom_point()` for a scatterplot.")
  }
  pass("Excellent! Now the plot clearly distinguishes both schools by color and shape.")
})
```

### Step 4c: Visualising Group Differences in Study Outcomes

```{r cp4-final-plot-mcq, echo=FALSE}
question("What is the main benefit of mapping both color and shape to school?",
  answer("It allows us to easily distinguish between the two schools, even if the plot is printed in black-and-white.", correct = TRUE,
         message = "Exactly! Using both color and shape makes the plot accessible and clear."),
  answer("It changes the underlying correlation coefficient.",
         message = "No, changing plot aesthetics does not alter statistical calculations."),
  answer("It filters outliers from the data.",
         message = "Plot aesthetics don't change the data."),
  answer("It doubles the sample size.",
         message = "No, sample size is unchanged."),
  allow_retry = TRUE
)
```

### Congratulations, you completed all exercises for this chapter!
